clarify

### 2025-09-18 _dev/clarify_old.Rmd

---
title: "`clarify`: Simulation-Based Inference for Regression Models"
date: "`r Sys.Date()`"
abstract: >
  Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects and adjusted predictions. It avoids several assumptions required by the delta method that are commonly violated, especially in nonlinear models, but does not require models to be refit as when using the bootstrap. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form the sampling distribution of the quantity. `clarify` provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. `clarify` supports a large and growing number of models through its interface with the `marginaleffects` package and provides native support for multiply imputed data.
draft: true
author:  
  - name: Noah Greifer
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    url: ngreifer.github.io
    orcid: 0000-0003-3067-7154
    email:  ngreifer@iq.harvard.edu
  - name: Steven Worthington
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: sworthington@iq.harvard.edu
    orcid: 0000-0001-9550-5797
  - name: Stefano Iacus
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: siacus@iq.harvard.edu
    orcid: 0000-0002-4884-0047
  - name: Gary King
    affiliation: Harvard University
    address:
    - Institute for Quantitative Social Science
    - Cambridge, MA
    email: king@harvard.edu
    orcid: 0000-0002-5327-7631
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(
  fig.path = "figures/",
  fig.align='center',
  fig.height = 2
)

if (!requireNamespace("clarify")) {
  install.packages("clarify")
}

if (!requireNamespace("MatchIt")) {
  install.packages("MatchIt")
}

if (!requireNamespace("Amelia")) {
  install.packages("Amelia")
}
```

@@

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. The usual method for estimating the uncertainty of the derived quantities is known as the "delta method", which involves two approximations: 1) that the variance of the derived quantity can be represented as a first-order Taylor series, and 2) that the estimate of the derived quantity is normally distributed[^1]. In many cases, especially with nonlinear models, these approximation can fail badly. For example, predicted probabilities close to 0 or 1 or ratios of coefficients or probabilities typically do not have normal (or even symmetrical) distributions in finite samples, and the usual Wald-type confidence interval limits produced from delta method standard errors can include values outside the domain of the quantity of interest.

@@

\CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating the sampling distributions of the derived quantities. Simulation-based inference is not only often more accurate than using the delta method, it is also simpler to understand and implement, as it does not require understanding Taylor series or the calculus that underlies it. This makes it more palatable to nontechnical audiences and easier to learn for students without sacrificing statistical performance.

The methodology \CRANpkg{clarify} relies on is described in @kingMakingMostStatistical2000 and @raineyCarefulConsiderationCLARIFY2023. Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a "posterior" distribution, which is taken to be the sampling distribution of the derived quantity, and from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates. The delta method makes these same assumptions in addition to the approximations about the variance and the shape of the sampling distribution of the estimates of the derived quantities.

The nonparametric bootstrap is another alternative to the delta method for inference that does not require these approximations; it typically involves re-sampling individuals from the sample, fitting the model in each sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters.

More formally, we fit a regression model $y_i = f(x_i; \beta)$, such as a linear or other generalized linear model with model coefficients $\beta$. We assume$$
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
$$where $\hat{\beta}$ is the vector of estimates of $\beta$ and $\Sigma_{\hat{\beta}}$ is their asymptotic covariance matrix. We define a function $\tau(\beta)$ that represents a quantity of interest derived from the model parameters, and compute its estimate $\widehat{\tau(\beta)}$ as $\tau(\hat{\beta})$. Next, we take $M$ draws $\tilde{\beta}^{(j)}$ for $j\in(1, \dots, M)$ from a multivariate Normal distribution with mean vector $\mu = \hat{\beta}$ and covariance $\Sigma = \hat{\Sigma}_{\hat{\beta}}$, where $\hat{\Sigma}_{\hat{\beta}}$ is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of $\tilde{\tau}=\tau(\tilde{\theta})$ as the "posterior" sampling distribution of $\widehat{\tau(\beta)}$, and compute its variance as$$
\sigma^2_{\widehat{\tau(\theta)}} = \frac{1}{M}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
$$and quantile $100(1-\alpha)\%$ confidence interval limits as $\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]$ where $\tilde{\tau}_{(k)}$ is the $k$th value of $\tilde{\tau}$ when arranged in ascending order.

Similar functionality exists in the \pkg{CLARIFY} package in Stata[^2] [@tomzClarifySoftwareInterpreting2003] and used to be available in the `Zelig` R package [@imaiCommonFrameworkStatistical2008a], though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

[^2]: Despite the similar name, the R package \CRANpkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \CRANpkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the "simulation-induced bias" described by @raineyCarefulConsiderationCLARIFY2023.

## Using \CRANpkg{clarify}

There are four steps to using \CRANpkg{clarify}:

1.  Fit the model to the data using modeling functions in supported packages.

2.  Use `sim()` to take draws from the sampling distribution of the estimated model coefficients.

3.  Use `sim_apply()` or its wrappers `sim_setx()`, `sim_ame()`, and `sim_adrf()` to compute derived quantities using each simulated set of coefficients.

4.  Use `summary()` and `plot()` to summarize and visualize the distribution of the derived quantities and perform inference on them.

In the sections below, we will describe how to implement these steps in detail. First, we will load \CRANpkg{clarify} using `library()`.

```{r}
library(clarify)
```

For a running example, we will use the `lalonde` dataset in the \CRANpkg{MatchIt} package [@JSSv042i08], which contains data on 614 participants enrolled in a job training program or sampled from a survey [@dehejiaCausalEffectsNonexperimental1999]. The treatment variable is `treat` and the outcome is `re78`, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of `treat` on `re78`, we will use it more generally to demonstrate all of \CRANpkg{clarify}'s capabilities. In addition, we will use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

### 1. Fitting the model

The first step is to fit the model. \CRANpkg{clarify} can operate on a large set of models (those supported by \CRANpkg{marginaleffects}), including generalized linear models, multinomial models, multivariate models, and instrumental variable models, many of which are available in other R packages. Even if \CRANpkg{clarify} does not offer direct support for a given model, there are ways to use its functionality regardless (explained in more detail below).

Because we are computing derived quantities, it is not critical to parameterize the model in such a way that the coefficients are interpretable, e.g., by using a model with interpretable coefficients or centering predictors. Below, we will fit a probit regression model for the outcome given the treatment and confounders. Coefficients in probit regression do not have a straightforward interpretation, but that does not matter; our quantities of interest can be expressed as derived quantities---functions of the model parameters, such as predictions, counterfactual predictions, and averages and contrasts of them.

@@

### 2. Drawing from the coefficient distribution

After fitting the model, we will use `sim()` to draw coefficients from their sampling distribution. The sampling distribution is assumed to be multivariate Normal or multivariate T with appropriate degrees of freedom, with a mean vector equal to the estimated coefficients and a covariance matrix equal to the asymptotic covariance matrix extracted from the model. The arguments to `sim()` are listed below:

@@

-   `vcov` -- either the covariance matrix of the estimated coefficients, a function used to extract it from the model (e.g., `sandwich::vcovHC()` for the robust covariance matrix), or a string or formula giving a code for extracting the covariance matrix, which is passed to `marginaleffects::get_vcov()`. If left unspecified, the default covariance matrix will be extracted from the model.

@@

-   `dist` -- the name of the distribution from which to draw the sampled coefficients. Can be `"normal"` for a Normal distribution or `t(#)` for a t-distribution, where `#` represents the degrees of freedom. If left unspecified, `sim()` will decide on which distribution makes sense given the characteristics of the model (the decision is made by `insight::get_df()` with `type = "wald"`). Typically this does not need to be specified.

If one's model is not supported by \CRANpkg{clarify}, one can omit the `fit` argument and just specify the `vcov` and `coefs` argument, which will draw the coefficients from the distribution named in `dist` (`"normal"` by default).

@@

# Drawing 1000 simulated coefficients using an HC2 robust
# covariance matrix
s <- sim(fit, n = 1000,
         vcov = "HC2")

@@

### 3. Computing derived quantities

After sampling the coefficients, we will compute derived quantities on each set of sampled coefficients and store the result, which represents the "posterior" distribution of the derived quantity, as well as on the original coefficients, which are used as the final estimates. The core functionality is provided by `sim_apply()`, which accepts a `clarify_sim` object from `sim()` and a function to compute and return one or more derived quantities, then applies that function to each set of simulated coefficients. The arguments to `sim_apply()` are below:

@@

The `FUN` argument can be specified in one of two ways: either as a function that takes in a model fit object (e.g., a `glm` or `lm` object, the output of a call to `glm()` or `lm()`) or a function that takes in a vector of coefficients. The latter will always work but the former only works for supported models. When the function takes in a model fit object, `sim_apply()` will first insert each set of sampled coefficients into the model fit object and then supply the modified model to `FUN`.

For example, we will let our derived quantity of interest be the predicted probability of the outcome for participant PSID1. We would specified our `FUN` function as follows:

@@

The resulting `clarify_est` object contains the simulated estimates in matrix form as well as the estimate computed on the original coefficients. We will examine the sampling distribution shortly, but first we will demonstrate computing a derived quantity from the coefficients directly.

The `race` variable is a factor, and the `black` category is used as the reference level, so it is not immediately clear whether there is a difference between the coefficients `racehispan` and `racewhite`, which represent the non-reference categories `hispan` and `white`. To compare these two directly, we can use `sim_apply()` to compute a derived quantity that corresponds to the difference between them.

@@

The function supplied to `FUN` can be arbitrarily complicated and return as many derived quantities as one wants, though the slower each run of `FUN` is, the longer it will take to simulate the derived quantities. Using parallel processing by supplying an argument to `cl` can sometimes dramatically speed up evaluation.

There are several functions in \CRANpkg{clarify} that serve as convenience wrappers for `sim_apply()` to automate some common derived quantities of interest. These include

@@

-   `reference` -- whether to display a normal density over the plot for each estimate. The default is `FALSE` to omit it.

@@

```{r, fig.width=4}
plot(est1, reference = TRUE, ci = FALSE)
```

Overlaid on the plot in red is a Normal distribution with the same mean and standard deviation as the simulated values; this is requested by setting `reference = TRUE`. From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds are clearly more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The plot itself is a `ggplot` object that can be modified using `ggplot2` syntax.

@@

-   `null` -- an optional argument specifying the null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if `method` is `"wald"`) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.

Inverting the uncertainty interval involves finding the smallest confidence level such that the null value is within the confidence bounds. The p-value for the test is one minus this level. It is only valid as a p-value when the simulated distribution of the estimates differs from its true sampling distribution under the null value by a location shift (which can be violated when the distribution of the estimates is asymmetric). However, because the p-values are invariant to monotonic transformations of the estimates and null value, as long as the distribution of *some* monotonic transformation of the estimate is a location shift from its sampling distribution under the null hypothesis, the p-values will be valid.

We can use `summary()` with the default arguments on our first `clarify_est` object to view the point estimate and quantile-based uncertainty bounds.

@@

A benefit of quantile-based confidence bounds is that the bounds will always be in the domain of the quantity of interest, e.g., bounds for probabilities will not be above 1 or below 0 if the model always generates predicted probabilities between 0 and 1. In contrast, Wald-type symmetric confidence intervals (including those produced by the delta method) can include values outside the domain of the quantity of interest. While it is possible to compute Wald-type confidence interval bounds that respect the domain boundaries, e.g., by computing the confidence interval around the logit of the predicted probability and using the inverse logit of the bounds as the bound on the probability, quantile confidence intervals do not require the user to know which transformation of the quantity of interest will yield appropriate bounds.

Our second estimated quantity, the difference between two regression coefficients, is closer to Normally distributed, as the plot below demonstrates (and would be expected theoretically), so we will use the Normal approximation to test the hypothesis that difference differs from 0.

```{r, fig.width=4}
plot(est2, reference = TRUE, ci = FALSE)

@@

The uncertainty intervals and p-values in the `summary()` output are computed using the Normal approximation because we set `method = "wald"`, and the p-value for the test that our estimate is equal to 0 is returned because we set `null = 0`. Note that the Normal approximation should be used only when the simulated sampling distribution is both close to Normally distributed and centered around the estimate (i.e., when the mean of the simulated values [red vertical line] coincides with the estimate computed on the original coefficients [black vertical line]).

## `sim_apply()` Wrappers: `sim_setx()`, `sim_ame()`, `sim_adrf()`

`sim_apply()` can be used to compute the simulated sampling distribution for an arbitrary derived quantity of interest, but there are some quantities that are common in applied research and may otherwise be somewhat challenging to program by hand, so \CRANpkg{clarify} provides shortcut functions to make computing these quantities simple. These functions include `sim_setx()`, `sim_ame()`, and `sim_adrf()`. Each of these can be used only when regression models compatible with \CRANpkg{clarify} are supplied to the original call to `sim()`.

@@

### `sim_setx()`: predictions at representative values

`sim_setx()` provides an interface to compute predictions at representative and user-supplied values of the predictors. For example, we might want to know what the effect of treatment is for a "typical" individual, which corresponds to the contrast between two model-based predictions (i.e., one under treatment and one under control for a unit with "typical" covariate values). This functionality mirrors the `setx()` and `setx1()` functionality of `Zelig` (which is where its name originates) and provides similar functionality to functions in \CRANpkg{modelbased}, \CRANpkg{emmeans}, \CRANpkg{effects}, and \CRANpkg{ggeffects}.

For each predictor, the user can specify whether they want predictions at specific values or at "typical" values, which are defined in \CRANpkg{clarify} as the mode for unordered categorical and binary variables, the median for ordered categorical variables, and the mean for continuous variables. Predictions for multiple predictor combinations can be requested by specifying values that will be used to create a grid of predictor values, or the grid itself can be supplied as a data frame of desired predictor profiles. In addition, the "first difference", defined here as the difference between predictions for two predictor combinations, can be computed.

@@

Here, we will use `sim_setx()` to examine predicted values of the outcome for control and treated units, at `re75` set to 0 and 20000, and `race` set to "black".

@@

We can plot the distributions of the simulated values using `plot()`, which also separates the predictions by the predictor values (it is often clearer without the uncertainty bounds). The `var` argument controls which variable is used for faceting the plots.

```{r, fig.width=5}
plot(est3, var = "re75", ci = FALSE)
```

One can see again how a delta method or Normal approximation may not have yielded valid uncertainty intervals given the non-Normality of the distributions.

@@

```{r, fig.width=5}
plot(est4)
```

We will return to display average dose-response functions using `sim_adrf()` later.

@@

### `sim_ame()`: average adjusted predictions and average marginal effects

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units[^3] [@longRegressionModelsCategorical2014].

[^3]: In \CRANpkg{marginaleffects}, AAPs are computed using `avg_predictions()`, AMEs for binary variables are computed using `avg_comparisons()`, and AMEs for continuous variables are computed using `avg_slopes()`. AAPs are sometimes known as average "counterfactual" predictions.

@@

-   `contrast` -- the name of an effect measure used to contrast AAPs. For continuous outcomes, `"diff"` requests the difference in means, but others are available for binary outcomes, including `"rr"` for the risk ratio, `"or"` for the odds ratio, and `"nnt"` for the number needed to treat, among others. If not specified, only AAPs will be computed if the variable named in `var` is categorical or specific values of the focal variable are specified in `var`. Ignored when the variable named in `var` is continuous and no specific values are specified because the AME is the only quantity computed. When `var` names a multi-category categorical variable, `contrast` cannot be used; see the section describing `transform()` for computing contrasts with them.

@@

-   `type` -- a string containing the type of predicted value to return. In most cases, this can be left unspecified to request predictions on the scale of the outcome (e.g., probabilities for binary outcomes).

@@

Here, we will use `sim_ame()` to compute the AME of `treat` just among those who were treated (in causal inference, this is known as the average treatment effect in the treated, or ATT [@greiferChoosingCausalEstimand2023]). We will request our estimate to be on the risk ratio scale.

```{r}
est6 <- sim_ame(s,
                var = "treat",
                subset = treat == 1,
                contrast = "rr",
                verbose = FALSE)
```

We can use `summary()` to display the estimates and their uncertainty intervals. Here, we will also use `null` to include a test for the null hypothesis that the risk ratio is equal to 1.

@@

We can compute the AME for a continuous predictor. Here, we will consider `age` (just for demonstration; this analysis does not have a valid interpretation).

@@

The AME is named `E[dY/d(age)]`, which signifies that a derivative has been computed (more precisely, the average of the unit-specific derivatives). This estimate can be interpreted like a slope in a linear regression model, but as a single summary of the effect of a predictor it is often too coarse to capture nonlinear relationships. The section below explains how to compute average dose-response functions for continuous predictors, which provide a more complete picture of their effects on an outcome.

Below, we will examine effect modification of the ATT by the predictor `married` using the `by` argument to estimate AAPs and their ratio within levels of `married`:

```{r}
est6b <- sim_ame(s,
                 var = "treat",
                 subset = treat == 1,
                 by = ~married,
                 contrast = "rr",
                 verbose = FALSE)

summary(est6b)
```

@@

``` r
sim_adrf(sim = , var = , subset = , by = , contrast = , at = ,
         n = , outcome = , type = , eps = , verbose = , cl = )
```

@@

Here, we will consider `age` (just for demonstration; this analysis does not have a valid interpretation) and compute the ADRF and AMEF of `age` on the outcome. We will only examine ages between 18 and 50, even though the range of `age` goes slightly beyond these values. First, we will compute the ADRF of `age`, which examines how the outcome would vary on average if one set all unit's value of `age` to each value between 18 and 50 (here we only use even ages to speed up computation).

@@

```{r, fig.width=5}
plot(est8)
```

From the plot, we can see that as `age` increases, the expected outcome decreases.

@@

Next we will compute the AMEF, the effect of `age` at each level of `age`.

@@

```{r, fig.width=5}
plot(est9)
```

>>>>

---
title: "clarify"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{clarify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6.5,
  fig.height = 2.75,
  fig.align = "center"
)
```

@@

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. The usual method for estimating the uncertainty of the derived quantities is an approximation known as the "delta method". The delta method involves two approximations: 1) that the variance of the derived quantity can be represented as a first-order Taylor series, and 2) that the resulting estimate is normally distributed[^1].In many cases, especially with nonlinear models, these approximation can fail badly. `clarify` implements an alternative to the delta method---simulation-based inference---which involves simulating the sampling distributions of the derived quantities.

@@

The methodology `clarify` relies on is described in @kingMakingMostStatistical2000 and @raineyCarefulConsiderationCLARIFY2023. Similar functionality exists in the `CLARIFY` package in Stata [@tomzClarifySoftwareInterpreting2003][^2] and used to be available in the `Zelig` R package [@imaiCommonFrameworkStatistical2008a], though there are differences in these implementations. `clarify` provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). `clarify` relies on and can be seen as a companion to the `marginaleffects` package, which offers similar functionality but primarily uses the delta method for calculating uncertainty.

[^2]: Despite the similar name, the R package `clarify` and the Stata package `CLARIFY` differ in several ways, one of which is that the estimates reported by `clarify` in R are those computed using the original model coefficients, whereas those reported by `CLARIFY` in Stata are those computed as the average of the simulated distribution. The R implementation avoids the "simulation-induced bias" described by @raineyCarefulConsiderationCLARIFY2023.

## Using `clarify`

There are four steps to using `clarify`:

1.  Fit the model to the data using modeling functions in supported packages

2.  Use `sim()` to take draws from the sampling distribution of the estimated model coefficients

3.  Use `sim_apply()` or its wrappers `sim_setx()`, `sim_ame()`, and `sim_adrf()` to compute derived quantities using each simulated set of coefficients

4.  Use `summary()` and `plot()` to summarize and visualize the distribution of the derived quantities and perform inference on them

In the sections below, we'll describe how to implement these steps in detail. First, we'll load `clarify` using `library()`.

```{r setup}
library(clarify)
```

For a running example, we'll use the `lalonde` dataset in the `MatchIt` package, which contains data on 614 participants enrolled in a job training program or sampled from a survey. The treatment variable is `treat` and the outcome is `re78`, and all other variables are confounders. Although the original intent was to estimate the effect of `treat` on `re78`, we'll use it more generally to demonstrate all of `clarify`'s capabilities. In addition, we'll use a transformation of the outcome variable to demonstrate applications to nonlinear models.

@@

## 1. Fitting the model

The first step is to fit the model. `clarify` can operate on a large set of models (those supported by `marginaleffects`), including generalized linear models, multinomial models, multivariate models, and instrumental variable models, many of which are available in other R packages. Even if `clarify` does not offer direct support for a given model, there are ways to use its functionality regardless (explained in more detail below).

Because we are computing derived quantities, it is not critical to parameterize the model in such a way that the coefficients are interpretable. Below, we'll fit a probit regression model for the outcome given the treatment and confounders. Coefficients in probit regression do not have a straightforward interpretation, but that's okay; our quantities of interest can be expressed as derived quantities--functions of the model parameters, such as predictions, counterfactual predictions, and averages and contrasts of them.

@@

## 2. Drawing from the coefficient distribution

After fitting the model, we will use `sim()` to draw coefficients from their sampling distribution. The sampling distribution is assumed to be multivariate normal or multivariate t with appropriate degrees of freedom, with a mean vector equal to the coefficient vector and a covariance matrix equal to the asymptotic covariance matrix extracted from the model. The arguments to `sim()` are listed below:

@@

-   `vcov` -- either the covariance matrix of the estimated coefficients, a function used to extract it from the model (e.g., `sandwich::vcovHC()` for the robust covariance matrix), or a string or formula giving a code for extracting the covariance matrix (see `marginaleffects::get_vcov()` for details). If left unspecified, the default covariance matrix will be extracted from the model.

@@

-   `dist` -- the name of the distribution from which to draw the sampled coefficients. Can be `"normal"` for a normal distribution or `t(#)` for a t-distribution, where `#` represents the degrees of freedom. If left unspecified, `sim()` will decide on which distribution makes sense given the characteristics of the model (the decision is made by `insight::get_df()` with `type = "wald"`). Typically this does not need to be specified.

If your model is not supported by `clarify`, you can omit the `fit` argument and just specify the `vcov` and `coefs` argument, which will draw the coefficients from the distribution named in `dist` (`"normal"` by default).

@@

# Drawing simulated coefficients using an HC2 robust
# covariance matrix
s <- sim(fit, vcov = "HC2")

@@

## 3. Computing derived quantities

After sampling the coefficients, we will compute derived quantities on each set of sampled coefficients and store the result, which represents a "posterior" distribution of the derived quantity, as well as on the original coefficients, which are used as the final estimates. The core functionality is provided by `sim_apply()`, which accepts a `clarify_sim` object from `sim()` and a function to compute and return one or more derived quantities, then applies that function to each set of simulated coefficients. The arguments to `sim_apply()` are below:

@@

The `FUN` argument can be specified in one of two ways: either as a function that takes in a model fit object or a function that takes in a vector of coefficients. The latter will always work but the former only works for supported models. When the function takes in a model fit object, `sim_apply()` will first insert each set of sampled coefficients into the model fit object and then supply the modified model to `FUN`.

For example, let's say our derived quantity of interest is the predicted probability of the outcome for participant PSID1. We would specified our `FUN` function as follows:

@@

The resulting `clarify_est` object contains the simulated estimates in matrix form as well as the estimate computed on the original coefficients. We'll examine the sampling distribution shortly, but first we'll demonstrate computing a derived quantity from the coefficients directly.

The `race` variable is a factor, and the `black` category is used as the reference level, so it's not immediately clear whether there is a difference between the coefficients `racehispan` and `racewhite`, which represent the non-reference categories `hispan` and `white`. To compare these two directly, we can use `sim_apply()` to compute a derived quantity that corresponds to the difference between them.

@@

The function supplied to `FUN` can be arbitrarily complicated and return as many derived quantities as you want, though the slower each run of `FUN` is, the longer it will take to simulate the derived quantities. Using parallel processing by supplying an argument to `cl` can sometimes dramatically speed up evaluation.

There are several functions in `clarify` that serve as convenience wrappers for `sim_apply()` to automate some common derived quantities of interest. These include

@@

-   `reference` -- whether to display a normal density over the plot for each estimate. The default is `FALSE` to omit them.

@@

```{r}
plot(est1, reference = TRUE, ci = FALSE)
```

From the plot, we can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling below 0 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds are clearly more appropriate, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The plot itself is a `ggplot` object that can be modified using `ggplot2` syntax.

@@

-   `null` -- an optional argument specifying the null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if `method` is `"quantile"`) or an inversion of the uncertainty interval. The default is not to display any p-values.

Inverting the uncertainty interval involves finding the smallest confidence level such that the null value is within the confidence bounds. The p-value for the test is one minus this level. It is only valid as a p-value when the simulated distribution of the estimates differs from its true sampling distribution under the null value by a location shift (which can be violated when the distribution of the estimates is asymmetric). However, because the p-values are invariant to monotonic transformations of the estimates and null value, as long as the distribution of some monotonic transformation of the estimate is a location shift from its sampling distribution under the null hypothesis, the p-values will be valid.

We'll use `summary()` with the default arguments on our first `clarify_est` object to view the point estimate and quantile-based uncertainty bounds.

@@

Our second estimated quantity, the difference between two regression coefficients, is closer to Normally distributed, as the plot below demonstrates (and would be expected theoretically), so we'll use the Normal approximation to test the hypothesis that difference differs from 0.

```{r}
plot(est2, reference = TRUE, ci = FALSE)

@@

The uncertainty intervals and p-values in the `summary()` output are computed using the Normal approximation because we set `method = "wald"`, and the p-value for the test that our estimate is equal to 0 is returned because we set `null = 0`. The computed bounds are very close to the quantile-based bound (displayed in the plot, and can be requested in `summary()` by setting `method = "quantile"`) because the simulated sampling distribution is close to Normal. Note that the Normal approximation should be used only when the simulated sampling distribution is both close to Normally distributed and centered around the estimate (i.e., when the mean of the simulated values [red vertical line] coincides with the estimate computed on the original coefficients [black vertical line]).

# `sim_apply()` wrappers: `sim_setx()`, `sim_ame()`, `sim_adrf()`

`sim_apply()` can be used to compute the simulated sampling distribution for any arbitrary derived quantity of interest, but there are some quantities that are common in applied research and may otherwise be somewhat challenging to program on their own, so `clarify` provides shortcut functions to make computing these quantities simple. These functions include `sim_setx()`, `sim_ame()`, and `sim_adrf()`. Each of these can only be used when regression models compatible with `clarify` are supplied to the original call to `sim()`.

@@

## `sim_setx()`: predictions at representative values

`sim_setx()` provides an interface to compute predictions at representative and user-supplied values of the predictors. For example, we might want to know what the effect of treatment is for a "typical" individual, which corresponds to the contrast between two model-based predictions (i.e., one under treatment and one under control for a unit with "typical" covariate values). This functionality mirrors the `setx()` and `setx1()` functionality of `Zelig` (which is where its name originates) and provides similar functionality to functions in `modelbased`, `emmeans`, `effects`, and `ggeffects`.

For each predictor, the user can specify whether they want predictions at specific values or at "typical" values, which are defined in `clarify` as the mode for unordered categorical and binary variables, the median for ordered categorical variables, and the mean for continuous variables. Predictions for multiple predictor combinations can be requested by specifying values that will be used to create a grid of predictor values, or the grid itself can be supplied as a data frame of desired predictor profiles. In addition, the "first difference", defined here as the difference between predictions for two predictor combinations, can be computed.

@@

Here, we'll use `sim_setx()` to examine predicted values of the outcome for control and treated units, at `re75` set to 0 and 20000, and `race` set to "black".

@@

We can plot the distributions of the simulated values using `plot()`, which also separates the predictions by the predictor values (it's often clearer without the uncertainty bounds). The `var` argument controls which variable is used for faceting the plots.

```{r}
plot(est3, var = "re75", ci = FALSE)
```

We can see again how a delta method or Normal approximation may not have yielded valid uncertainty intervals given the non-Normality of the distributions.

@@

```{r}
plot(est4)
```

We'll return to display average dose-response functions using `sim_adrf()` later.

@@

## `sim_ame()`: average adjusted predictions and average marginal effects

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs) and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units[^3].

[^3]: In `marginaleffects`, AAPs are computed using `avg_predictions()`, AMEs for binary variables are computed using `avg_comparisons()`, and AMEs for continuous variables are computed using `avg_slopes()`. AAPs are sometimes known as average "counterfactual" predictions.

@@

-   `contrast` -- the name of an effect measure used to contrast AAPs. For continuous outcomes, `"diff"` requests the difference in means, but others are available for binary outcomes, including `"rr"` for the risk ratio, `"or"` for the odds ratio, and `"nnt"` for the number needed to treat, among others. If not specified, only AAPs will be computed if the variable named in `var` is binary. Ignored when the variable named in `var` is continuous because the AME is the only quantity computed. When `var` names a multi-category categorical variable, `contrast` cannot be used; see the section describing `transfom()` for computing contrasts with them.

@@

@@

Here, we'll use `sim_ame()` to compute the AME of `treat` just among those who were treated (in causal inference, this is known as the average treatment effect in the treated, or ATT). We'll request our estimate to be on the risk ratio scale.

```{r}
est6 <- sim_ame(s,
                var = "treat", subset = treat == 1,
                contrast = "RR", verbose = FALSE)
```

We can use `summary()` to display the estimates and their uncertainty intervals. Here, we'll also use `null` to include a test for the null hypothesis that the risk ratio is equal to 1.

@@

We can compute the AME for a continuous predictor. Here, we'll consider `age` (just for demonstration; this analysis does not have a valid interpretation).

@@

The AME is named `E[dY/d(age)]`, which signifies that a derivative has been computed (more precisely, the average of the unit-specific derivatives). This estimate can be interpreted like a slope in a linear regression model, but as a single summary of the effect of a predictor it is too coarse to capture nonlinear relationships. The section below explains how to compute average dose-response functions for continuous predictors, which provide a more complete picture of their effects on an outcome.

Below, we'll examine effect modification using the `by` argument to estimate AAPs and their ratio within levels of the predictor `married`:

```{r}
est6b <- sim_ame(s,
                 var = "treat", by = ~married,
                 contrast = "RR", verbose = FALSE)

est6b
```

@@

``` r
sim_adrf(sim = , var = , subset = , contrast = , at = ,
         n = , outcome = , type = , eps = , verbose = ,
         cl = )
```

@@

Here, we'll consider `age` (just for demonstration; this analysis does not have a valid interpretation) and compute the ADRF and AMEF of `age` on the outcome. We'll only examine ages between 18 and 50, even though the range of `age` goes slightly beyond these values. First, we'll compute the ADRF of `age`, which examines how the outcome would vary on average if one set all unit's value of `age` to each value between 18 and 50 (here we only use even ages to speed up computation).

@@

```{r}
plot(est8)
```

From the plot, we can see that as `age` increases, the expected outcome also increases.

@@

Next we'll compute the AMEF, the effect of `age` at each level of `age`.

@@

```{r}
plot(est9)
```
